An Open Repository of Real-Time COVID-19 Indicators

### 2021-09-27 indicators/code/indicator-paper-figures.Rmd
New backfill figure

```{r setup, message=FALSE}
library(covidcast)
library(gridExtra)
library(ggplot2)
library(scales)
library(dplyr)
library(purrr)
```

@@

@@

```{r figure-6-redux, message=FALSE, cache=TRUE}
cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv") %>%
  filter(Date == "07/20/2021")

>>>>

```{r setup, message=FALSE}
library(covidcast)
library(gridExtra)
library(ggplot2)
library(scales)
library(dplyr)
library(tidyr)
library(purrr)
```

@@

Amount of backfill of key signals from various sources, for comparison:

```{r figure-6, message=FALSE, cache=TRUE}
SOURCES <- c("chng", "fb-survey", "jhu-csse", "jhu-csse")
SIGNALS <- c("smoothed_outpatient_cli", "smoothed_whh_cmnty_cli",
             "confirmed_incidence_prop", "deaths_incidence_prop")
NAMES <- c("lag", "CHNG-CLI", "CTIS-CLI-in-community", "Cases", "Deaths")

calc_pct_abs_diff <- function(source, signal, start_date, end_date,
                              final_lag, lags) {
  pctile_abs_change <- numeric(length(lags))

  dates <- seq.Date(as.Date(start_date),
                    as.Date(end_date),
                    by = "1 day")

  # finalized value
  df_final <- covidcast_signal(source, signal, start_date, end_date,
                               as_of = end_date + final_lag,
                               geo_type = "state") %>%
    select(geo_value, time_value, value)

  for (ii in seq_along(lags)) {
    lag <- lags[ii]
    errs <- c()

    for (jj in seq_along(dates)) {
      date <- dates[jj]

      df_prelim <- covidcast_signal(source, signal, date, date,
                                    as_of = date + lag,
                                    geo_type = "state")
      if (nrow(df_prelim) == 0) {
        next
      }

      df_prelim <- df_prelim %>%
        select(geo_value, time_value, value) %>%
        mutate(lag = lag)

      errs_this_date <- df_final %>%
        inner_join(df_prelim,
                   by = c("geo_value", "time_value"),
                   suffix = c(".final", ".prelim")) %>%
        mutate(abs_diff = abs(value.prelim - value.final) / value.final) %>%
        pull(abs_diff)

      errs <- c(errs, errs_this_date)
    }

    pctile_abs_change[ii] <- quantile(errs, probs = 0.95, na.rm = TRUE,
                                      names = FALSE)
  }

  return(pctile_abs_change)
}

format_df <- function(df) {
  names(df) <- NAMES

  df <- df %>%
    pivot_longer(!lag, names_to = "signal", values_to = "value")

  return(df)
}

start_date <- as.Date("2020-10-15")
end_date <- as.Date("2021-04-15")
final_lag <- 120
lags <- c(90, 70, 50, 30, 10)

pctile_df <- data.frame(lag = lags)

for (ii in seq_along(SOURCES)) {
  source <- SOURCES[ii]
  signal <- SIGNALS[ii]

  pctile_df[[signal]] <- calc_pct_abs_diff(source, signal, start_date, end_date,
                                           final_lag, lags)
}

pctile_df <- format_df(pctile_df)

pctile_df %>%
  ggplot(aes(x = lag, y = value, group = signal, color = signal)) +
  geom_line() +
  geom_text(data = filter(pctile_df, lag == min(lags)),
            aes(x = lag, y = value, label = signal, hjust = "left"),
            inherit.aes = FALSE, nudge_y = 0.015) +
  labs(x = "Days since initial issue",
       y = "95th percentile of absolute percent error",
       color = "Signal") +
  scale_y_continuous(labels = label_percent()) +
  scale_x_continuous(breaks = lags) +
  theme_bw() +
  theme(legend.position = "none")

ggsave("../paper/fig/backfill-compare.pdf", width = 5, height = 3.5)
```

## Figure 7

@@

```{r figure-7, message=FALSE, cache=TRUE}
cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv") %>%
  filter(Date == "07/20/2021")

### 2021-09-27 indicators/paper/paper.tex
New backfill figure

Claims-based signals typically undergo heavy backfill as additional claims are
processed and errors are corrected; the median relative error between initial
reports and final values is over 20\% for such data, and only after roughly 35
days do estimates typically match finalized values within 5\%. However, the
systematic nature of this backfill, as illustrated in Figure~\ref{fig:dv_as_of},
suggests that statistical models could be fit (potentially separately for each
location) to estimate the final values from preliminary reports. On the other
hand, official public health reporting of COVID cases and deaths can be subject
to revision as death certificates are audited and backlogs cleared, resulting in
thousands of cases and deaths being added or removed. This process is much more
difficult to predict, and thus claims data and other sources may be a useful
stand-in while public health reports are aggregated and corrected.

To reiterate a previous point, when training and validating forecast models (on
historical data), users will want to use data that was known \textit{as of} the
forecast date, not revised versions that only became available much later. The
COVIDcast API makes all historical versions available and easily accessible for
this purpose; and this feature plays a prominent role in our own analysis of
forecasting and hotspot prediction models appearing in a companion paper.

>>>>

Claims-based signals typically undergo heavy backfill as additional claims are
processed and errors are corrected; the median relative error between initial
reports and final values is over 20\% for such data, and only after roughly 35
days do estimates typically match finalized values within 5\%. However, the
systematic nature of this backfill, as illustrated in Figure~\ref{fig:dv_as_of},
suggests that statistical models could be fit (potentially separately for each
location) to estimate the final values from preliminary reports. On the other
hand, official public health reporting of COVID cases and deaths can be subject
to revision as death certificates are audited and backlogs cleared, resulting in
thousands of cases and deaths being added or removed.
Figure~\ref{fig:backfill-compare} shows the amount of revision for several
signals, in terms of the 95th percentile of relative change between initial
estimates and revised estimates issued later; this shows that reported deaths
can have large revisions comparable to revisions in claims-based signals. This
backfill is much more difficult to predict, and thus claims data and other
sources may be a useful stand-in while public health reports are aggregated and
corrected.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
  \caption{95th percentile relative changes between initial values of key
    signals, compared to the final values reported much later. Calculated using
    each day's state-level estimates for the six months between October 15, 2020
    and April 15, 2021; final values were those reported 120 days after the end
    of this period. Even officially reported case and death data can have large
    revisions 30--60 days after initial reporting; much of this is driven by
    individual large revisions affecting specific states and dates, rather than
    systematic changes affecting all states and dates.}
  \label{fig:backfill-compare}
\end{figure}

To reiterate a previous point, when training and validating forecast models that
use signals (on historical data), users will want to use data that was known
\textit{as of} the forecast date, not revised versions that only became
available much later. The COVIDcast API makes all historical versions available
and easily accessible for this purpose; and this feature plays a prominent role
in our own analysis of forecasting and hotspot prediction models appearing in a
companion paper. Revisions to forecast targets, such as cases and deaths, also
complicate the evaluation of forecast models: comparisons of forecasts to the
ground truth may shift when the ground truth is revised weeks later. Only by
systematically tracking revisions can this effect be monitored.
